### 2022-08-20 paper/paper.tex
Write 3.1

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2022

% ready for submission
% \usepackage{neurips_2022}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
% \usepackage[preprint]{neurips_2022}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{neurips_2022}
\bibliographystyle{plain}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2022}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}

\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\mb}[1]{\mathbf{#1}}
\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\def\EE{\mathbb{E}}
\def\PP{\mathbb{P}}
\def\QQ{\mathbb{Q}}
\def\II{\mathbb{I}}
\def\CC{\mathbb{C}}
\def\NN{\mathbb{N}}
\def\OO{\mathbb{O}}
\def\HH{\mathbb{H}}
\def\Ll{\mathcal{L}}
\def\Mm{\mathcal{M}}
\DeclareMathOperator*{\argmin}{arg\,min}

\title{Cross-Dataset Recommender System for \\Overcoming Selection Bias}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Fengyu Li \\
  Cornell University\\
  \texttt{fl334@cornell.edu} \\
  \And
  Sarah Dean \\
  Cornell University \\
  \texttt{sdean@cornell.edu} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph. The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Introduction}
Selection bias is one of the most prevalent sources of biases for recommender systems \cite{chen2020bias}. Selection bias happens when there is a pattern in the users' ratings that is unique to the training set. For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' tastes \cite{pradel2012ranking}. However, the environment the recommender system is deployed on contains all movies regardless of personal tastes. This discrepancy produces a misalignment between training and deployed settings, which is known as a distribution shift. Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms\cite{schnabel2016recommendations}.

When recommender systems are deployed in real-world platforms, it is arguably likely that the clients are able to collect rating-associated data from different source distributions. These data (or feedbacks) are either implicit or explicit \cite{aggarwal2016recommender}, while implicit data often comes in a highly-quantized, binary form and explicit data is usually less quantized. For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data. This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and also users' innate biases.

In this paper, we attempt to take advantage of datasets from differently quantized sources. More specifically, we proposed a way to feed both a 5-quantized dataset and a binary dataset to any gradient-based recommender algorithm.  To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a single distribution. We design experiments under a simulated environment and shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.

Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical} and thus is more suitable for training, we decided to use it as the training data of matrix factorization and use the more-quantized dataset for propensity scoring and deriving the inverse-probability-scoring (IPS) estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms. In this way, our cross-dataset learning framework empowers existing recommender algorithms to make use of the more quantized, less biased data. We carried out experiment and found our method outperforming baselines by a significant margin.

% Quantization is a practical way to categorize an input from a continuous distribution to a discrete set of bins. We assume the ground truth of a user's rating is a continuous distribution over items because of the inherent impossibility of numerically expressing the feeling of enjoyment and excitement. We treat a recommender dataset as a noised quantization of the ground truth. In other words, users convert their feelings into one of the discrete choices when they rate an item. Recommender datasets are quantized into different degrees. For example, a typical movie rating dataset uses a 5-quantization. A tinder like/dislike dataset uses a 2-quantization.

\section{Related Work}
Prior works on overcoming selection bias-induced distribution shift via a propensity-based approach begins with the seminal paper \cite{schnabel2016recommendations}, which introduces the IPS method into recommender systems. Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach. In this work we add to previous attempts the idea of using a more quantized, implicit dataset for more accurate propensity estimation.

The IPS method in our context is, essentially, a method of weighting training examples to correct the bias in the training data. Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate} \cite{zhang2018importance}. Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited.

\section{Susceptibility to Selection Bias}
We first examine the susceptibility to selection bias of differently quantized data by manually introducing biased distributions of various degrees to the differently quantized training sets. It is crucial for us that the 2-quantized datasets do not exhibit particular weakness when facing selection bias so that they can be properly adopted for propensity estimation.

\subsection{Simulated Environment for Controlling Bias}
Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets, which shall be explained in section 5. In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias. For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled. The effect of $\beta$ on probability of being observed for different ratings is visualized in xxx. In our experiment we assume constantly $10\%$ of ratings are observed.
\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

\subsection{Results}

\section{Cross-Dataset Propensity Estimation}

\[ \argmin_{V,W,A} \frac{1}{N}\left(\frac{(Y_{u,i}- (V_u^TW_i + A))^2}{P_{u,i}} \right) + c \Vert A \Vert^2\]
where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $\hat{Y} = V_u^TW_i + A$ is the predicted rating, and $c \Vert A \Vert^2$ is the regularizer.

\section{Experiment}
We designed experiments under simulated environments that verify the results introduced above.

\begin{table}
  \caption{Test set RMSE and MAE for MF-NBPE and baselines}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Datasets}
We used two datasets in our simulated environment:\\

\textbf{Imputed ML100K Dataset.}

\textbf{Latent Factors Simulated Dataset.}

\subsection{Baselines}

\subsection{Results}

\section{Conclusion}

\section*{Acknowledgements}

\bibliography{refs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}